---
title: "seurate workflow"
author: "Boyu Jiang"
date: '2022-07-08'
output: html_document
---
## load packages

```{r}
suppressMessages(library(seurate))
suppressMessages(library(seurate))
library(stringr)
library(dplyr)
```

## load data and check it
## using raw counts is the easiest way to process data through Seurat
```{r}
counts <- read.table("GSE111229_Mammary_Tumor_fibroblasts_768samples_rawCounts.txt.gz")
dim(counts)
fivenum(apply(counts,1,function(x) sum(x>0) ))
boxplot(apply(counts,1,function(x) sum(x>0) ))
fivenum(apply(counts,2,function(x) sum(x>0) ))
hist(apply(counts,2,function(x) sum(x>0) ))

dat=log2(edgeR::cpm(counts)+1) 
hc=hclust(dist(t(dat)))  
plot(hc,labels = FALSE)  
clus = cutree(hc, 4) 
group_list= as.factor(clus) 
table(group_list) 
```

## extract batch information
## splite the colnames by '_', and get the thrid column
## here we define the genes with more than 1 read as expressed, and count their number in each sample
```{r}
plate=str_split(colnames(dat),'_',simplify = T)[,3]
table(plate)
n_g = apply(counts,2,function(x) sum(x>1))
```
## create dataframe that includes the attributes of cells
```{r}
df=data.frame(g=group_list,plate=plate,n_g=n_g)
meta=df
identical(rownames(meta), colnames(counts))
```
 
## create seurat object(V3)
## min.cell--gene must eb expressed in at least 5 cells
## min.genes-- gene must have at least 1000 reads
```{r}
sce <- CreateSeuratObject(counts  = counts,
                          meta.data = meta,
                          min.cells = 5,
                          min.genes = 2000, 
                          project = "sce")
```

## visualization of feature
```{r}
plot1 <- VlnPlot(object = sce,
                 features = c("nFeature_RNA", "nCount_RNA"),
                 group.by = 'plate',
                 ncol = 2)
plot2 <- VlnPlot(object = sce, 
                 features = c("nFeature_RNA", "nCount_RNA"), 
                 group.by = 'g',
                 ncol = 2)
CombinePlots(plots = list(plot1, plot2))
```

## add ERCC or MT to sce for later QC
```{r}
sce[["percent.ercc"]] <- PercentageFeatureSet(sce, pattern = "^ERCC-")
VlnPlot(object = sce, 
        features = c("nFeature_RNA", "nCount_RNA", "percent.ercc" ), 
        group.by = 'g',
        ncol = 3)
```
## normalization
```{r}
sce <- NormalizeData(object = sce,
                     normalization.method = "LogNormalize",
                     scale.factor = 10000)
sce[['RNA']][1:3,1:3]
```
## identify HVGs（highly variable genes）
```{r}
sce <- FindVariableFeatures(sce, selection.method = "vst")
VariableFeatures(sce)
```
## scaling 
## result is stored in sce@assays$RNA@scale.data
```{r}
all.genes <- rownames(sce)
length(rownames(sce))
sce <- ScaleData(sce, features = all.genes,
                 vars.to.regress = c("nCount_RNA","percent.ercc" ))
```


## dimensionality reduction--PCA
## result is stored in sce@reductions$pca@feature.loadings
```{r}
sce <- RunPCA(sce, 
              features = VariableFeatures(object = sce),
              ndims.print = 1:5, 
              nfeatures.print = 5)
DimPlot(sce, reduction = "pca",group.by = 'plate')
DimPlot(sce, reduction = "pca",group.by = 'g')
ElbowPlot(sce)
```
## clustering
## the higher resolution, the more number of clusters,when set resolution as 0.4, we get number of communities: 4
```{r}
sce <- FindNeighbors(sce, dims = 1:15)
sce <- FindClusters(sce, resolution = 0.4)
```

## another way 0f dimensionality reduction--tSNE
```{r}
sce <- RunTSNE(object = sce, dims.use = 1:15, do.fast = TRUE)
DimPlot(sce,reduction = "tsne",label=T)
```
## find marker genes
## min.pct: the percent of a gene in any two gouprs of cells must higher than a value
```{r}
cluster2.markers <- FindMarkers(sce, ident.1 = 1, min.pct = 0.25)
head(cluster2.markers, n = 5)
markers_genes =  rownames(head(cluster2.markers, n = 5))
VlnPlot(sce, features = markers_genes)
FeaturePlot(sce, features = markers_genes)
```

## find all markers and make heatmap
```{r}
sce.markers <- FindAllMarkers(sce, only.pos = TRUE, 
                              min.pct = 0.25, 
                              logfc.threshold = 0.25)
top20 <- sce.markers %>% group_by(cluster) %>% top_n(n = 20, wt = avg_log2FC)
DoHeatmap(sce, features = top20$gene) + NoLegend()
```

